---
title: "Classifier Tutorial"
output: html_notebook
---

### Load the Source Code
```{r}
source("~/Desktop/Morris Lab/CellTag-Classifier/src/Classifier_Source.R")
```

### Load the Example Dataset (Before Starcode Collapsing) and Look at All CellTags
```{r}
# load celltag counts (before Starcode collapse)
count.celltag <- read.csv("~/Desktop/Morris Lab/CellTag-Classifier/data/2-tag species mixing before collapsing.csv", header = TRUE, row.names=1)
# all CellTags, pre-filtering
all.celltags <- rowSums(count.celltag)
hist(all.celltags, breaks = 50)
```

### Get the Starcode Collapsed Results and Collapse the CellTag Count Matrix
#### Note: Please refer to here for more information about Starcode: https://github.com/gui11aume/starcode
#### Options used for Starcode: -s --print-clusters
```{r}
# load starcode collapsed result
starcode.out <- read.table("~/Desktop/Morris Lab/CellTag-Classifier/data/2-tag species mixing starcode result.txt", sep = "\t", stringsAsFactors=FALSE)

# create a list to store consensus celltags for each centroid celltag
consensus.lookup <- list()
for (i in 1:nrow(starcode.out)) { # repeat for each row
  centroid <- 0
  # split consensus celltags separated by comma and populate list
  centroid <- strsplit(starcode.out[i,3], ",")
  names(centroid) <- starcode.out[i,1]
  consensus.lookup[[i]] <- split(unname(centroid), names(centroid))
}

# create new matrix where each count.celltag column is renamed to its centroid CellTag
count.celltag.collapsed <- as.data.frame(count.celltag)
for (j in 1:ncol(count.celltag.collapsed)) {
  n <- grep(colnames(count.celltag.collapsed[j]), consensus.lookup)
  colnames(count.celltag.collapsed)[j] <- names(consensus.lookup[[n]])
}

# collapse count.celltag by consensus CellTags
count.celltag.collapsed <- t(count.celltag.collapsed)
count.celltag.collapsed <- by(count.celltag.collapsed, INDICES=row.names(count.celltag.collapsed), FUN=colSums)
count.celltag.collapsed <- as.data.frame(do.call(cbind,count.celltag.collapsed))
```

### First Round of Normalization (Normalizing Between Cells)
```{r}
# Log the count matrix
count.norm.expr <- log2(count.celltag.collapsed+1)
ct.dge.norm <- t(count.norm.expr)
rownames(ct.dge.norm) <- colnames(count.celltag.collapsed)
colnames(ct.dge.norm) <- rownames(count.celltag.collapsed)
# Normalize
norm.ct.dge <- normalize.function(ct.dge.norm)
```

### Pull Out the Most Abundant CellTags
```{r}
count.norm.expr.t <- t(norm.ct.dge)

# # pull out dge by 5 most abundant celltags
ct.dge <- count.norm.expr.t[,c("TGCTATAT", "GTTGGCTA")]
collapsed.orig.count <- count.celltag.collapsed[,c("TGCTATAT", "GTTGGCTA")]
```

### Second Round of Normalization (Normalizing Between CellTags)
```{r}
norm.ct.dge.2 <- normalize.function(ct.dge)
```

### Dynamic CellTag Detection
```{r}
perc.ls <- dynamic.celltag.detection(norm.ct.dge.2)
```

### Binarization and Classification
```{r}
bin.class.ct <- binarization.classification(norm.ct.dge.2, perc.ls)
```

### Check the Binary Result
```{r}
hist(rowSums(bin.class.ct))
table(rowSums(bin.class.ct))
```

### CellTag Calling
```{r}
bin.class.ct <- multiplet.calling(bin.class.ct)
# check results
barplot(table(bin.class.ct$ct.call))
```

### Multiplet Secondary Check
```{r}
bin.class.ct <- multiplet.checkpoint(bin.class.ct, perc.ls, multiplet.table.path = "~/Desktop/Morris Lab/CellTag-Classifier/data/Expected Multiplet Rate.csv")
```

### Human/Mouse Assignment
#### Note: convert celltag calls to human/mouse calls: pSmal B (GTTGGCTA) was mouse, all other tags were human
```{r}
mouse.human.tag <- data.frame(row.names = c("Mouse", "Human"), Tag = c("GTTGGCTA", "TGCTATAT"))
bin.class.ct$ct.call.2 <- bin.class.ct$ct.call
bin.class.ct[which(bin.class.ct$ct.call == mouse.human.tag["Mouse", "Tag"]),"ct.call.2"] <- "mouse"
bin.class.ct[which(bin.class.ct$ct.call == mouse.human.tag["Human", "Tag"]),"ct.call.2"] <- "human"
```

### Compare CellTag Classification with 10x Results
```{r}
# load 10x classification of human, mouse, and doublet cells
class.10x <- read.table("~/Desktop/Morris Lab/CellTag-Classifier/data/2-tag species mixing 10x gem_classification.csv", sep = ",", header = TRUE)
rownames(class.10x) <- gsub('-.*$', "", class.10x$barcode)
colnames(class.10x)[4] <- "10x.call"

# remove nd cells and compare with 10x
class.ct.filt <- subset(bin.class.ct, ct.call.2!="nd")
class.10x.filt <- class.10x[row.names(class.ct.filt),]

# creat composite of human and mouse classification, by 10x and CellTag
classification <- merge(class.10x.filt$`10x.call`, class.ct.filt$ct.call.2, by=0)
classification <- classification[,2:3]
colnames(classification) <- c("10x", "CellTag")
table(classification)
```

### Calculate Cohen's Kappa Score
```{r}
suppressMessages(library(psych))
rslt <- cohen.kappa(matrix(table(classification), 3, 3))
print(paste0("Cohen's Kappa Score = ", rslt$kappa))
```

